---
title: "Final Exam - Rissler Energy Analysis"
author: "Alex Rhomberg"
date: "12/12/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggformula)
load("Weather.Rda")
load("Data.Rda")
```

## House Energy Trends

```{r Usage by Month}
FromAlliantConsumptionRaw %>% filter(STARTTIME != is.na(STARTTIME)) %>% group_by(year, month) %>% 
  summarize(Usage = sum(USAGE)) %>% gf_point(Usage ~ month| year) %>% gf_labs(x="Month of Year", y = "Usage")
```

```{r Weather by Month}
WeatherRaw %>% group_by(year, month) %>% summarize(temp = mean(apparentTemperature)) %>% gf_point(temp ~ month | year) %>% gf_labs(x="Month of Year", y = "Apparent Temp Avg")
```

The first graph above shows monthly energy usage KWH over the span of 2018 and 2019. The second graph shows the apparent temperature F over the same duration. At first look, there does not seem to be much relation between the two graphs, but it can be seen that when weather spikes at a high or a low, energy usage is almost always at a high peak, showing more extreme temperatures resulted in greater energy usage.

```{r Usage by days of the month for each month}
FromAlliantConsumptionRaw %>% filter(STARTTIME != is.na(STARTTIME)) %>% group_by(month, day) %>% 
  summarize(Usage = sum(USAGE)) %>% gf_point(Usage ~ day | month) %>% gf_labs(x="Day of Month", y = "Usage")
```

```{r Weather by days of the month for each month}
WeatherRaw %>% group_by(month, day) %>% summarise(temp = mean(apparentTemperature)) %>% gf_point(temp ~ day | month) %>% gf_labs(x="Month", y = "Apparent Temp Avg")
```

Here we have a similar display of energy usage KWH and apparent temperature F over the days of the month for the 12 months. Most of the energy usage stays the same throughout the year until the warm summer months are hit, then the energy usage increases greatly. The respective months in the second graph show very consistently high tempertatures, alluding that most probably Rissler does not like his house to be warm in the summer.

```{r Usage by hour of the day}
FromAlliantConsumptionRaw %>% filter(STARTTIME != is.na(STARTTIME)) %>% group_by(hour, minute) %>% 
  summarize(MonthUsage = sum(USAGE)) %>% gf_point(MonthUsage ~ hour ) %>% gf_labs(x="Hour", y = "Usage")
```

```{r Weather by hour of the day}
WeatherRaw %>% group_by(hour, minute) %>% summarise(temp = mean(apparentTemperature)) %>% gf_point(temp ~ hour) %>% gf_labs(x="Hour", y = "Apparent Temp Avg")
```

These two graphs show the energy usage HWK and apparent temperature F over the 24 hours in a day. The graphs follow similar curves, but are off set from each other in time. Energy usage begins to increase as temperature starts to decrease from its peak and vice versa.

```{r Usage of each hour of the day for each day of the month}
FromAlliantConsumptionRaw %>% filter(STARTTIME != is.na(STARTTIME)) %>% group_by(day, hour) %>% 
  summarize(MonthUsage = sum(USAGE)) %>% gf_point(MonthUsage ~ hour | day) %>% gf_labs(x="Hour", y = "Usage")
```

```{r Weather of each hour of the day for each day of the month}
WeatherRaw %>% group_by(day, hour) %>% summarise(temp = mean(apparentTemperature)) %>% gf_point(temp ~ hour | day) %>% gf_labs(x="Hour", y = "Apparent Temp Avg")
```

These two graphs show the energy usage KWH and apparent temperature F over the hours of the day with each day of the month marked. These graphs follow similar trends as the one above, it is just interesting to see how various days throughout the month differ from each other. 

## Electric Heat Pump Water Heater Switch

```{r Usage before Water Heater}
Rissler_955_Kirkwood_St %>% filter(WaterHeater == 0) %>% gf_line(TotalUsage ~ Month, color = "Red") %>% gf_line(Bill ~ Month, color = "Green") %>% gf_theme(axis.text.x = element_text(angle = 90, hjust = 1)) %>% gf_labs(x="Time", y = "Usage")
```

```{r Usage After Heater}
Rissler_955_Kirkwood_St %>% filter(WaterHeater == 1) %>% gf_line(TotalUsage ~ Month, color = "Red") %>% gf_line(Bill ~ Month, color = "Green") %>% gf_theme(axis.text.x = element_text(angle = 90, hjust = 1)) %>% gf_labs(x="Time", y = "Usage")
```

```{r Average Usage After Heater}
Rissler_955_Kirkwood_St %>% group_by(WaterHeater) %>% summarise(UsageMean = mean(TotalUsage), BillMean = mean(Bill))
```


In short, making the switch to an electric heat pump water heater was benneficial. The first graph shows the energy usage KWH in Red over time and the bill price in Green, and the second graph shows the same thing but after the electric heat pump water heater was installed. The energy usage before installing sits around the 300 - 600 KWH usage with a bill average close to 100. After installation, the usage range drops to 200 - 400 on average with a bill well below 100, averaging around 25 - 50.

## Enphase vs Alliant

```{r Usage by Month vs Enphase}
FromAlliantConsumptionRaw %>% filter(STARTTIME != is.na(STARTTIME), year == 2019) %>% group_by(year, month) %>% 
  summarize(Usage = sum(USAGE)) %>% gf_point(Usage ~ month| year) %>% gf_labs(x="Month of Year", y = "Usage")
```

```{r Enphase Usage by Month}
EnphaseRaw %>% filter(DATETIME != is.na(DATETIME), year == 2019) %>% group_by(year, month) %>% summarize(Usage = sum(EnergyConsumed)) %>% gf_point(Usage ~ month | year) %>% gf_labs(x="Month of Year", y = "Usage")
```

The first graph above is from the first analysis of this document and shows the energy usage KWH of the months of the year with the data from Alliant. The second graph shows the energy usage KWH from the enphase device on Rissler's house. It can be seen that the two follow almost identical trends, meaning Alliant is producing accurate records in allign with the enphase device. 

```{r Usage by hour of the day vs Enphase}
FromAlliantConsumptionRaw %>% filter(STARTTIME != is.na(STARTTIME)) %>% group_by(hour, minute) %>% 
  summarize(MonthUsage = sum(USAGE)) %>% gf_point(MonthUsage ~ hour ) %>% gf_labs(x="Hour", y = "Usage")
```

Similarly, the first graph from above is taken from the inital analysis on this document and shows energy usage KWH over the hours of the day with information from Alliant. The second graph shows the same thing but from the enphase device. Both graphs follow similar trends but are not exactly accurate, showing there could be a margin of error with the enphase device, or that the information from Alliant is wrong. 

```{r Enphase usage by hour of the day}
EnphaseRaw %>% filter(DATETIME != is.na(DATETIME)) %>% group_by(hour, minute) %>% summarize(MonthUsage = sum(EnergyConsumed)) %>% gf_point(MonthUsage ~ hour) %>% gf_labs(x="Hour", y = "Usage")
```


## Consistency of Bills

```{r Bill vs Total Usage vs Grid Usage vs Grid Production}
Rissler_955_Kirkwood_St %>% gf_line(Bill ~ Month, color = "Purple") %>% gf_line(TotalUsage ~ Month, color = "Blue") %>% gf_line(UsageGrid ~ Month, color = "Red") %>% gf_line(ProductionToGrid ~ Month, color = "Green") %>% gf_labs(x="Multivariate", y = "Time")
```

```{r Weather Overview vs Bill}
WeatherRaw2 %>% gf_line(apparentTemperature ~ time) %>% gf_labs(x="Time", y = "Temperature")
```

The first graph above shows many different aspects of the bill and energy usages over time. The Red line shows energy usage KWH with a respective purple line representing the bill price. The blue line shows the usage from the grid after solar and water heat pump installation with the green line showing solar production to the grid. The second graph shows the apparent temperatures during the same duration of time.

Before the installations in 2018, it is easy to follow the trends of the the energy usage and the price of the bill. They remain consistent to each other for the most part, showing similar humps and dips. It is a little harder to read after the installations. Thinking of it numerically, adding the red and green lines would equal the blue line. When looking at the bill, if the green line is lower, the bill should be higher and vice versa, which can be seen as true in the plot. Overall, the bill seems to be consistent with the energery usage ammounts reported by Alliant. When comparing to weather, it is interesting to see that when the weather drops well below 0, it is almost a gaurantee the energy bill and usage will spike high same with temperatures above 90, a similar trend that can be seen in the first couple of graphs on this ananlysis document. 


## Time of Day Pricing, Good?

```{r Usage by hour of the day 2}
FromAlliantConsumptionRaw %>% filter(STARTTIME != is.na(STARTTIME)) %>% group_by(hour, minute) %>% 
  summarize(MonthUsage = sum(USAGE)) %>% gf_point(MonthUsage ~ hour ) %>% gf_labs(x="Hour", y = "Usage")
```

The graph above shows the monthly usage on the y axis with the hour of the day on the x axis.

Most of the energy usage takes place from 3:00 pm to 6:00 am. With the time of day prices during the hours of 8:00 pm to 7:00 am being close to half off in both seasons, switching to this plan makes the most amount of sense. Seventy Five percent of the high energy usage takes place when the price of usage is close to half off, so a significant amount of money is saved because of this. Basically, a 50% coupon for those times. Making this switch is a smart decision if most of your energy usage falls between the discounted hours, not vice versa.
