---
title: "Final Melt Presentation"
author: "Alex, Gio, Arias"
date: "11/15/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggformula)
library(readxl)
library(readr)
library(lubridate)
load("CleanData.rda")
load("CleanWeatherData.rda")
```

## MELT 2016

```{r MELT 2016}
clean2016 %>%  group_by(flg_NetDep, Dropped) %>% 
  count() %>% spread(Dropped, n)
```
326/336 is who stayed = 0.9702381

## Apparent Temp 2016

```{r Apparent Temp 2016}
cleanWeather2016 %>% gf_point(apparentTemperature ~ HOUR, color = ~ month(time)) %>% gf_smooth(apparentTemperature~HOUR, method = loess, colorse = TRUE) %>% gf_facet_wrap(~month(time))
```

## Humidity 2016 A

```{r Humidity 2016a}
cleanWeather2016 %>% gf_point(humidity ~ HOUR, color = ~ month(time)) %>% gf_smooth(humidity~HOUR, method = loess, colorse = TRUE) %>% gf_facet_wrap(~month(time))
```

## Humidity 2016 B

```{r Humidity 2016b}
cleanWeather2016 %>% gf_point(humidity ~ windSpeed, color = ~ month(time)) %>% gf_smooth(humidity~windSpeed, method = loess, colorse = TRUE) %>% gf_facet_wrap(~month(time))
```

## 2016 Overview

- Second most amount of people left (10).
- Apparent temperature from 10:00AM to 08:00PM was above 90F.
- Many nights with high humidity.
- Little wind throughout the days.


## MELT 2017

```{r MELT 2017}
clean2017 %>%  group_by(flg_NetDep, Dropped) %>% 
  count() %>% spread(Dropped, n)
```
332/339 is who stayed = 0.97935103

## Apparent Temp 2017

```{r Apparent Temp 2017}
cleanWeather2017 %>% gf_point(apparentTemperature ~ HOUR, color = ~ month(time)) %>% gf_smooth(apparentTemperature~HOUR, method = loess, colorse = TRUE) %>% gf_facet_wrap(~month(time))
```

## Humidity 2017 A

```{r Humidity 2017a}
cleanWeather2017 %>% gf_point(humidity ~ HOUR, color = ~ month(time)) %>% gf_smooth(humidity~HOUR, method = loess, colorse = TRUE) %>% gf_facet_wrap(~month(time))
```

## Humidity 2017 B

```{r Humidity 2017b}
cleanWeather2017 %>% gf_point(humidity ~ windSpeed, color = ~ month(time)) %>% gf_smooth(humidity~windSpeed, method = loess, colorse = TRUE) %>% gf_facet_wrap(~month(time))
```


## 2017 Overview

- Only 7 people decided to drop.
- Apparent temperatures ranged from 46F to 87F in the two months.
- Many nights and mornings with high humidity.
- Light wind throughout the days.



## MELT 2018

```{r MELT 2018}
clean2018 %>%  group_by(flg_NetDep, Dropped) %>% 
  count() %>% spread(Dropped, n)
```
356/368 is who stayed = 0.9673913

## Apparent Temp 2018

```{r Apparent Temp 2018}
cleanWeather2018 %>% gf_point(apparentTemperature ~ HOUR, color = ~ month(time)) %>% gf_smooth(apparentTemperature~HOUR, method = loess, colorse = TRUE) %>% gf_facet_wrap(~month(time))
```

## Humidity 2018 A

```{r Humidity 2018a}
cleanWeather2018 %>% gf_point(humidity ~ HOUR, color = ~ month(time)) %>% gf_smooth(humidity~HOUR, method = loess, colorse = TRUE) %>% gf_facet_wrap(~month(time))
```

## Humidity 2018 B

```{r Humidity 2018b}
cleanWeather2018 %>% gf_point(humidity ~ windSpeed, color = ~ month(time)) %>% gf_smooth(humidity~windSpeed, method = loess, colorse = TRUE) %>% gf_facet_wrap(~month(time))
```


## 2018 Overview

- The most amount of people left (12).
- High appaarent tempertatures.
- Humidity was above 80% most days.
- Wind was below 10 most days.


## MELT 2019

```{r MELT 2019}
clean2019 %>%  group_by(flg_NetDep, Dropped) %>% 
  count() %>% spread(Dropped, n)
```
324/330 is who stayed = 0.98181818

## Apparent Temp 2019

```{r Apparent Temp 2019}
cleanWeather2019 %>% gf_point(apparentTemperature ~ HOUR, color = ~ month(time)) %>% gf_smooth(apparentTemperature~HOUR, method = loess, colorse = TRUE) %>% gf_facet_wrap(~month(time))
```

## Humidity 2019 A

```{r Humidity 2019a}
cleanWeather2019 %>% gf_point(humidity ~ HOUR, color = ~ month(time)) %>% gf_smooth(humidity~HOUR, method = loess, colorse = TRUE) %>% gf_facet_wrap(~month(time))
```

## Humidity 2019 B

```{r Humidity 2019b}
cleanWeather2019 %>% gf_point(humidity ~ windSpeed, color = ~ month(time)) %>% gf_smooth(humidity~windSpeed, method = loess, colorse = TRUE) %>% gf_facet_wrap(~month(time))
```


## 2019 Overview

- Only 6 people decided to drop.
- Apparent temperatures during the day ranged from 70F to 88F in the two months.
- The days saw low humidity.
- Humidity did not seem to have a huge impact.